Viscous Transport in Eroding Porous Media

### 2019-10-09 docs/transport/erosion_transport.tex
continued formatting

\usepackage{graphicx}
\usepackage{epstopdf, epsfig}

\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

@@

Flow in porous media is further complicated when boundaries dynamically
evolve in response to the fluid flow. This coupling between the geometry
and the flow occurs, for example, in applications involving
melting~\citep{bec-vis1998, rycroft2016asymmetric, jambon2018singular,
favier2019rayleigh, morrow2019moving},
dissolution~\citep{kan-zha-che-he2002, mac2015shape, moo2017,
wykes2018self}, deposition~\citep{joh-eli1995, hewett2018modelling},
biofilm growth~\citep{tan-val-wer2015}, and crack
formation~\citep{cho2019crack}. We focus on erosion, a fluid-mechanical
process that is prevalent in many geophysical, hydrological, and
industrial applications~\citep{berhanu2012shape, hewett2017evolution,
lachaussee2018competitive, lopez2018cfd, allen2019sde, amin2019role}.
When a porous geometry erodes, certain qualitative characteristics are
unveiled that affect transport through the geometry.  For example, an
eroded geometry may contain channels of high porosity, which, though few
in number and modest in volume fraction, transmit a large portion of the
flux \citep{qua-moo2018}. This arrangement results in velocities that
vary over several orders of magnitude~\citep{all-hea-lab-rei2002}.
Moreover, channelization creates heterogeneous and anisotropic medium
properties, which affect the transport of tracers such as
contaminants~\citep{cve-che-wen1996, dag1987, kon-bre1978} and
heat~\citep{nil-sto1990, ree-sto1995}.

@@

To compute stable simulations of erosion, we use high-order methods in
both space and time. The time integration method is unchanged from
previous work~\citep{qua-moo2018}.  We apply a mild regularization and a
smoothing term to eliminate numerical instabilities that can be triggered by changes in
sign of the shear stress, and we use a stable second-order Runge-Kutta
method applied to the {\thL} coordinates~\citep{hou-low-she1994} of the
eroding grains.  In this work, we introduce a new quadrature method to
resolve dense suspensions.  The accuracy of the
trapezoid rule, which was used in previous work, is adequate for
bodies that are sufficiently separated~\citep{tre-wei2014}.  However, it
loses accuracy whenever grains are in close contact.  One of the
earliest works to address quadrature methods for nearly-singular
integrands was developed by Baker and Shelley~\citep{bak-she1986}, and in
recent years, many other near-singular integration schemes have
followed~\citep{oja-tor2015, kli-tor2018, hel-oja2008a, bea-yin-wil2016,
bea-lai2001, klo-bar-gre-one2013}.  We use a Barycentric quadrature
method~\citep{bar2014, bar-wu-vee2015} since it is a non-intrusive
modification of the trapezoid rule, and the error is guaranteed to be
uniformly bounded.  We extend the original quadrature method to compute
the velocity gradient, which is needed to evaluate the shear stress and
the fluid vorticity.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Contributions}
Our first contribution is extending the Barycentric quadrature rule of
Barnett et al.~\citep{bar-wu-vee2015} to accurately compute velocity
gradients associated with Stokes flow. This new quadrature rule permits
accurate and stable simulations of erosion, and, moreover, enables
visualization of the vorticity field throughout.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Limitations}
As in our previous work~\citep{qua-moo2018}, we consider two-dimensional
immobile eroding grains. However, this work does not limit the proximity
of  grains. Compared to our previous work, we are able to simulate much
denser suspensions with a comparable number of discretization points and
without upsampling.  While we do investigate transport in this work, we
only consider passive particles that are advected by the flow.  To model
other physical processes, such as contaminant transport, we plan to
simulate a finite Peclet number advection-diffusion equation in eroded
geometries.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In
equations~\eqref{eqn:velocityDLP},~\eqref{eqn:deformationDLP},~\eqref{eqn:pressureDLP},
and~\eqref{eqn:vorticityDLP}, the fluid velocity, deformation tensor,
pressure, and vorticity are all written as layer potentials in $\RR^2$.
However, the quadrature method we introduce in Section~\ref{sec:method}
requires a complex-valued representation of the layer potentials for the
velocity, deformation tensor, and vorticity. The first step to form a
complex representation is to write the Laplace double-layer potential as
the complex integral
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{2\upi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2}\eeta(\yy)\, ds_\yy = \Re (v(x)) \quad 
    \text{where} \quad v(x) = \frac{1}{2\upi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy, \quad x \in \Omega,
  \label{eqn:laplaceComplex}
\end{align}
where $x = x_1 + i x_2,y = y_1 + i y_2 \in \CC$ are the complex
counterparts of $\xx = (x_1,x_2),\yy = (y_1,y_2) \in \RR^2$, and $\eta =
\eta_1 + i \eta_2$ is the complex counterpart of $\eeta =
(\eta_1,\eta_2)$. Therefore, depending on the formulation of the layer
potential, $\Omega$ is being interpreted as a subset of $\RR^2$ or
$\CC$. Equation~\eqref{eqn:laplaceComplex} is converted to a Cauchy
integral by first finding the boundary data of $v$.  Assuming that
$\Omega$ is a simply-connected interior domain, the boundary data of $v$
satisfies the Sokhotski-Plemelj jump relation
\begin{align}
  \label{eqn:SPrelation}
  v(x) = - \frac{1}{2} \eta(x) + \frac{1}{2\upi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \Omega.
\end{align}
For exterior domains, the jump term changes from $-1/2$ to $1/2$, and
for multiply-connected domains, such as a porous media, $\bd\Omega$ is
decomposed into its different connected components and the appropriate
jump relation is applied.  Having computed the boundary data of the
holomorphic function $v$, by the Cauchy integral theorem we have
\begin{align}
  \label{eqn:cauchy}
  v(x) = \frac{1}{2\upi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \quad
  v'(x) = \frac{1}{2\upi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \quad
  v''(x) = \frac{1}{\upi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy,
\end{align}
for $x \in \Omega$.  Since $v(x)$ depends on the complex-valued density
function $\eta$, we use the notation $v[\eta](x)$ for the holomorphic
function defined in equation~\eqref{eqn:laplaceComplex}, and its first
two derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.

Finally, the Stokes double-layer potential~\eqref{eqn:velocityDLP} can
be written using a Laplace double-layer
potential~\eqref{eqn:laplaceComplex} and its gradients 
\begin{equation}
  \label{eqn:Stokes2Laplace}
  \begin{aligned}
    \DDD[\eeta](\xx) &= 
      \frac{1}{2\upi} \int_{\bd\Omega} 
        \frac{\nn}{\rho^2} (\rr \cdot \eeta) \, ds_\yy + 
      \frac{1}{2\upi} \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2} (\yy \cdot \eeta) \, ds_\yy \\
      &- \frac{1}{2\upi} x_1 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_1(\yy) \, ds_\yy -
      \frac{1}{2\upi} x_2 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_2(\yy) \, ds_\yy.
  \end{aligned}
\end{equation}
Therefore, the Stokes double-layer potential can be written as a sum of
Cauchy integrals and its first derivative~\citep{bar-wu-vee2015}
\begin{equation}
  \begin{aligned}
    u_1(x) &= \Re (v[\psi_1](x)) + \Re (v'[y\cdot\eta](x)) 
             -x_1\Re (v'[\eta_1](x)) - x_2\Re (v'[\eta_2](x)), \\
    u_2(x) &= \Re (v[\psi_2](x)) - \Im (v'[y\cdot\eta](x)) 
         +x_1\Im (v'[\eta_1](x)) + x_2\Im (v'[\eta_2](x)),
  \end{aligned}
  \label{eqn:cauchyVelocity}
\end{equation}
where $y \cdot \eta = y_1 \eta_1 + i y_2 \eta_2 \in \CC$, 
\begin{align} 
  \psi_1=(\eta_1+i\eta_2)\frac{\Re(n)}{n}, \quad
  \psi_2=(\eta_1+i\eta_2)\frac{\Im(n)}{n},
\end{align}
and $n \in \CC$ is the complex counterpart of the outward unit normal
$\nn \in \RR^2$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation for the Gradient of the
Double-Layer Potential}
\label{sec:gradDLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Computing the shear stress and vorticity requires a complex-valued layer
potential representation of the velocity gradient.  The deformation
tensor at $x \in \Omega$ is found by computing the derivatives of the
expressions for $u_1$ and $u_2$ in equation~\eqref{eqn:cauchyVelocity}  
\begin{equation}
\label{eqn:cauchyGradient}
  \begin{aligned}
    \pderiv{u_1}{x_1} &= +\Re (v'[\psi_1](x)) + 
      \Re (v''[y\cdot\eta](x)) - \Re (v'[\eta_1](x)) - 
      x_1\Re (v''[\eta_1](x)) - x_2\Re (v''[\eta_2](x)), \\
    \pderiv{u_1}{x_2} &= - \Im (v'[\psi_1](x)) - 
      \Im (v''[y\cdot\eta](x)) + x_1\Im (v''[\eta_1](x)) - 
      \Re (v'[\eta_2](x)) + x_2\Im (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_1} &= +\Re (v'[\psi_2](x)) - 
      \Im (v''[y\cdot\eta](x)) + \Im (v'[\eta_1](x)) +
      x_1\Im (v''[\eta_1](x)) + x_2\Im (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_2} &= -\Im (v'[\psi_2](x)) - 
      \Re (v''[y\cdot\eta](x)) + x_1\Re (v''[\eta_1](x)) +
      \Im (v'[\eta_2](x)) + x_2\Re (v''[\eta_2](x)).
  \end{aligned}
\end{equation}
The same expressions are used to compute the deformation tensor for $x
\in \bd\Omega$, except that the jump
condition~\eqref{eqn:deformationJump} is included.  Finally, to compute
the shear stress, the deformation tensor on $\bd\Omega$ is applied to
the normal and tangent vectors as in equation~\eqref{eqn:shearStress}.
The velocity gradient is also used to compute the vorticity in the fluid
bulk.  For $x \in \Omega$, the Cauchy integral representation of the
vorticity is
\begin{align}
  \omega(x) = \pderiv{u_2}{x_1} - \pderiv{u_1}{x_2} = 
\Re (v'[\psi_2](x)) + \Im (v'[\psi_1](x)) 
 + \Re (v'[\eta_2](x))+ \Im (v'[\eta_1](x)), \quad x \in \Omega.
\end{align}

@@

There have been efforts to relate the tortuosity, $T$, to the porosity,
$\phi$.  For example, Matyka et al.~\citep{mat-kha-koz2008} propose the
models
\begin{subequations}
  \label{eqn:tortuosityModels}
  \begin{align}
    \widehat{T}(\phi) &= \phi^{-p}, \\
    \widehat{T}(\phi) &= 1-p \log \phi, \\
    \widehat{T}(\phi) &= 1+p (1-\phi), \\
    \widehat{T}(\phi) &= (1+p (1-\phi))^2, 
  \end{align}
\end{subequations}
where $p>0$ is a fitting parameter.  In Section~\ref{sec:results}, we
compare these four models with the tortuosity of eroding porous
geometries.

@@

We showed in Sections~\ref{sec:DLPcomplex} and~\ref{sec:gradDLPcomplex}
that the velocity, shear stress, and vorticity of the double-layer
representation can all be written as the sum of Cauchy integrals and its
first two derivatives.  Therefore, we can accurately compute the
velocity, shear stress, and vorticity at any target location by using a
quadrature rule with a uniform error bound for Cauchy integrals and its
derivatives~\eqref{eqn:cauchy}. Ioakimidis et al.~\citep{ioa-pap-per1991}
developed quadrature rules, that we call {\em Barycentric quadrature
rules}, to compute Cauchy integrals and their derivatives with an error
bound that is independent of $x \in \CC$.  Then, Barnett et
al.~\citep{bar-wu-vee2015} used these quadrature rules to compute the
Stokes double-layer potential representation of the
velocity~\eqref{eqn:velocityDLP}, and we briefly summarize their method.
Then, we extend their work to compute the second derivative so that the
shear stress and vorticity can be computed with a uniform error bound.

@@

To compute the shear stress and vorticity, we require a Barycentric
quadrature rule for $v''(x)$.  The derivation is largely based on the
work of Ioakimidis et al.~\citep{ioa-pap-per1991} (see equation (2.12)).
We start with the second derivative of the Cauchy integral theorem
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y)}{(y-x)^3}\,dy - v''(x).
\end{align}
For the interior case, $x \in \Omega$, we use the identity
\begin{align}
  \frac{1}{2\upi i}\int_{\bd\Omega} \frac{1}{(y-x)^n}\, dy = 
  \left\{
    \begin{array}{ll}
      1, & n = 1, \\
      0, & n = 2,3,\ldots. \\
    \end{array}
  \right.
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y) - v''(x)(y-x)^2 - 2v(x) - 2(y-x)v'(x)}
      {(y-x)^3}\, dy.
\end{align}
This integrand is constructed so that it is bounded for all $x \in
\Omega$, and applying the trapezoid rule, we have
\begin{align}
  0 \approx  \sum_{j=1}^{N} 
      \frac{2v^{-}(y_j) - v''(x)(y_j-x)^2 - 2v(x) - 2(y_j-x)v'(x)}
      {(y_j-x)^3} w_j,
\end{align}
where the accuracy is independent of $x$.  Solving for $v''(x)$, the
Barycentric quadrature rule for the interior second derivative is
\begin{align}
  v''(x) &\approx \frac{2\sum\limits_{j=1}^N 
    \frac{v^{-}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{1}{y_j-x}w_j}, \quad x \in \Omega.
\end{align}
For the exterior case, $x \in \Omega^c$, we start with the identity
\begin{align}
\frac{1}{x-a} &= -\frac{1}{2\upi i}\int_{\bd\Omega} 
    \frac{(y-a)^{-1}}{y-x}\, dy. 
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
    \frac{2v^+(y) - 2v(x) - 2(y-x)v'(x) + (y-a)^{-1} (x-a) (y-x)^2 v''(x)}
    {(y-x)^3}.
\end{align}
As in the interior case, the integrand is chosen so that it is bounded
for all $x \in \Omega^c$.  Therefore, after applying the trapezoid rule
and solving for $v''(x)$, we have the Barycentric quadrature rule for
the exterior second derivative
\begin{align}
  v''(x) &\approx \frac{1}{x-a}\frac{2\sum\limits_{j=1}^N
    \frac{v^{+}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j}, 
    \quad x \in \Omega^c.
\end{align}
The quadrature rule for $v''(x)$ requires the values $v(x)$ and $v'(x)$,
and theses are computed using the quadrature rules in
equations~\eqref{eqn:BaryvInterior},~\eqref{eqn:BaryvExterior},~\eqref{eqn:BaryvprimeInterior},
and~\eqref{eqn:BaryvprimeExterior}. In Section~\ref{sec:results}, these
quadrature rules are used to form simulations of nearly-touching eroding
grains, and to study dynamics of the flow in regions arbitrarily close
to eroding grains.

>>>>

@@

Flow in porous media is further complicated when boundaries dynamically
evolve in response to the fluid flow. This coupling between the geometry
and the flow occurs, for example, in applications involving
melting~\citep{bec-vis1998, rycroft2016asymmetric, jambon2018singular,
favier2019rayleigh, morrow2019moving},
dissolution~\citep{kan-zha-che-he2002, mac2015shape, moo2017,
wykes2018self}, deposition~\citep{joh-eli1995, hewett2018modelling},
biofilm growth~\citep{tan-val-wer2015}, and crack
formation~\citep{cho2019crack}. We focus on erosion, a fluid-mechanical
process that is prevalent in many geophysical, hydrological, and
industrial applications~\citep{berhanu2012shape, hewett2017evolution,
lachaussee2018competitive, lopez2018cfd, allen2019sde, amin2019role}.
When a porous geometry erodes, certain qualitative characteristics are
unveiled that affect transport through the geometry.  For example, an
eroded geometry may contain channels of high porosity, which, though few
in number and modest in volume fraction, transmit a large portion of the
flux~\citep{qua-moo2018}. This arrangement results in velocities that
vary over several orders of magnitude~\citep{all-hea-lab-rei2002}.
Moreover, channelization creates heterogeneous and anisotropic medium
properties, which affect the transport of tracers such as
contaminants~\citep{cve-che-wen1996, dag1987, kon-bre1978} and
heat~\citep{nil-sto1990, ree-sto1995}.

@@

To compute stable simulations of erosion, we use high-order methods in
both space and time. The time integration method is unchanged from
previous work~\citep{qua-moo2018}.  We apply a mild regularization and a
smoothing term to eliminate numerical instabilities that can be triggered by changes in
sign of the shear stress, and we use a stable second-order Runge-Kutta
method applied to the {\thL} coordinates~\citep{hou-low-she1994} of the
eroding grains.  In this work, we introduce a new quadrature method to
resolve dense suspensions.  The accuracy of the
trapezoid rule, which was used in previous work, is adequate for
bodies that are sufficiently separated~\citep{tre-wei2014}.  However, it
loses accuracy whenever grains are in close contact.  One of the
earliest works to address quadrature methods for nearly-singular
integrands was developed by~\citet{bak-she1986}, and in recent years,
many other near-singular integration schemes have
followed~\citep{oja-tor2015, kli-tor2018, hel-oja2008a, bea-yin-wil2016,
bea-lai2001, klo-bar-gre-one2013}.  We use a Barycentric quadrature
method~\citep{bar2014, bar-wu-vee2015} since it is a non-intrusive
modification of the trapezoid rule, and the error is guaranteed to be
uniformly bounded.  We extend the original quadrature method to compute
the velocity gradient, which is needed to evaluate the shear stress and
the fluid vorticity.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Contributions}
Our first contribution is extending the Barycentric quadrature rule
of~\citet{bar-wu-vee2015} to accurately compute velocity gradients
associated with Stokes flow. This new quadrature rule permits accurate
and stable simulations of erosion, and, moreover, enables visualization
of the vorticity field throughout.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Limitations}
As in our previous work~\citep{qua-moo2018}, we consider two-dimensional
immobile eroding grains. However, this work does not limit the proximity
of  grains. Compared to our previous work, we are able to simulate much
denser suspensions with a comparable number of discretization points and
without upsampling.  While we do investigate transport in this work, we
only consider passive particles that are advected by the flow.  To model
other physical processes, such as contaminant transport, we plan to
simulate a finite Peclet number advection-diffusion equation in eroded
geometries.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In
equations~\eqref{eqn:velocityDLP},~\eqref{eqn:deformationDLP},~\eqref{eqn:pressureDLP},
and~\eqref{eqn:vorticityDLP}, the fluid velocity, deformation tensor,
pressure, and vorticity are all written as layer potentials in $\RR^2$.
However, the quadrature method we introduce in Section~\ref{sec:method}
requires a complex-valued representation of the layer potentials for the
velocity, deformation tensor, and vorticity. The first step to form a
complex representation is to write the Laplace double-layer potential as
the complex integral
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{2\upi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2}\eeta(\yy)\, ds_\yy = \Real (v(x)) \quad 
    \text{where} \quad v(x) = \frac{1}{2\upi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy, \quad x \in \Omega,
  \label{eqn:laplaceComplex}
\end{align}
where $x = x_1 + i x_2,y = y_1 + i y_2 \in \CC$ are the complex
counterparts of $\xx = (x_1,x_2),\yy = (y_1,y_2) \in \RR^2$, and $\eta =
\eta_1 + i \eta_2$ is the complex counterpart of $\eeta =
(\eta_1,\eta_2)$. Therefore, depending on the formulation of the layer
potential, $\Omega$ is being interpreted as a subset of $\RR^2$ or
$\CC$. Equation~\eqref{eqn:laplaceComplex} is converted to a Cauchy
integral by first finding the boundary data of $v$.  Assuming that
$\Omega$ is a simply-connected interior domain, the boundary data of $v$
satisfies the Sokhotski-Plemelj jump relation
\begin{align}
  \label{eqn:SPrelation}
  v(x) = - \frac{1}{2} \eta(x) + \frac{1}{2\upi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \Omega.
\end{align}
For exterior domains, the jump term changes from $-1/2$ to $1/2$, and
for multiply-connected domains, such as a porous media, $\bd\Omega$ is
decomposed into its different connected components and the appropriate
jump relation is applied.  Having computed the boundary data of the
holomorphic function $v$, by the Cauchy integral theorem we have
\begin{align}
  \label{eqn:cauchy}
  v(x) = \frac{1}{2\upi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \quad
  v'(x) = \frac{1}{2\upi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \quad
  v''(x) = \frac{1}{\upi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy,
\end{align}
for $x \in \Omega$.  Since $v(x)$ depends on the complex-valued density
function $\eta$, we use the notation $v[\eta](x)$ for the holomorphic
function defined in equation~\eqref{eqn:laplaceComplex}, and its first
two derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.

Finally, the Stokes double-layer potential~\eqref{eqn:velocityDLP} can
be written using a Laplace double-layer
potential~\eqref{eqn:laplaceComplex} and its gradients 
\begin{equation}
  \label{eqn:Stokes2Laplace}
  \begin{aligned}
    \DDD[\eeta](\xx) &= 
      \frac{1}{2\upi} \int_{\bd\Omega} 
        \frac{\nn}{\rho^2} (\rr \cdot \eeta) \, ds_\yy + 
      \frac{1}{2\upi} \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2} (\yy \cdot \eeta) \, ds_\yy \\
      &- \frac{1}{2\upi} x_1 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_1(\yy) \, ds_\yy -
      \frac{1}{2\upi} x_2 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_2(\yy) \, ds_\yy.
  \end{aligned}
\end{equation}
Therefore, the Stokes double-layer potential can be written as a sum of
Cauchy integrals and its first derivative~\citep{bar-wu-vee2015}
\begin{equation}
  \begin{aligned}
    u_1(x) &= \Real (v[\psi_1](x)) + \Real (v'[y\cdot\eta](x)) 
             -x_1\Real (v'[\eta_1](x)) - x_2\Real (v'[\eta_2](x)), \\
    u_2(x) &= \Real (v[\psi_2](x)) - \Imag (v'[y\cdot\eta](x)) 
         +x_1\Imag (v'[\eta_1](x)) + x_2\Imag (v'[\eta_2](x)),
  \end{aligned}
  \label{eqn:cauchyVelocity}
\end{equation}
where $y \cdot \eta = y_1 \eta_1 + i y_2 \eta_2 \in \CC$, 
\begin{align} 
  \psi_1=(\eta_1+i\eta_2)\frac{\Real(n)}{n}, \quad
  \psi_2=(\eta_1+i\eta_2)\frac{\Imag(n)}{n},
\end{align}
and $n \in \CC$ is the complex counterpart of the outward unit normal
$\nn \in \RR^2$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation for the Gradient of the
Double-Layer Potential}
\label{sec:gradDLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Computing the shear stress and vorticity requires a complex-valued layer
potential representation of the velocity gradient.  The deformation
tensor at $x \in \Omega$ is found by computing the derivatives of the
expressions for $u_1$ and $u_2$ in equation~\eqref{eqn:cauchyVelocity}  
\begin{equation}
\label{eqn:cauchyGradient}
  \begin{aligned}
    \pderiv{u_1}{x_1} &= +\Real (v'[\psi_1](x)) + 
      \Real (v''[y\cdot\eta](x)) - \Real (v'[\eta_1](x)) - 
      x_1\Real (v''[\eta_1](x)) - x_2\Real (v''[\eta_2](x)), \\
    \pderiv{u_1}{x_2} &= - \Imag (v'[\psi_1](x)) - 
      \Imag (v''[y\cdot\eta](x)) + x_1\Imag (v''[\eta_1](x)) - 
      \Real (v'[\eta_2](x)) + x_2\Imag (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_1} &= +\Real (v'[\psi_2](x)) - 
      \Imag (v''[y\cdot\eta](x)) + \Imag (v'[\eta_1](x)) +
      x_1\Imag (v''[\eta_1](x)) + x_2\Imag (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_2} &= -\Imag (v'[\psi_2](x)) - 
      \Real (v''[y\cdot\eta](x)) + x_1\Real (v''[\eta_1](x)) +
      \Imag (v'[\eta_2](x)) + x_2\Real (v''[\eta_2](x)).
  \end{aligned}
\end{equation}
The same expressions are used to compute the deformation tensor for $x
\in \bd\Omega$, except that the jump
condition~\eqref{eqn:deformationJump} is included.  Finally, to compute
the shear stress, the deformation tensor on $\bd\Omega$ is applied to
the normal and tangent vectors as in equation~\eqref{eqn:shearStress}.
The velocity gradient is also used to compute the vorticity in the fluid
bulk.  For $x \in \Omega$, the Cauchy integral representation of the
vorticity is
\begin{align}
  \omega(x) = \pderiv{u_2}{x_1} - \pderiv{u_1}{x_2} = 
\Real (v'[\psi_2](x)) + \Imag (v'[\psi_1](x)) 
 + \Real (v'[\eta_2](x))+ \Imag (v'[\eta_1](x)), \quad x \in \Omega.
\end{align}

@@

There have been efforts to relate the tortuosity, $T$, to the porosity,
$\phi$.  For example,~\citet{mat-kha-koz2008} propose the
models
\begin{subequations}
  \label{eqn:tortuosityModels}
  \begin{align}
    \widehat{T}(\phi) &= \phi^{-p}, \\
    \widehat{T}(\phi) &= 1-p \log \phi, \\
    \widehat{T}(\phi) &= 1+p (1-\phi), \\
    \widehat{T}(\phi) &= (1+p (1-\phi))^2, 
  \end{align}
\end{subequations}
where $p>0$ is a fitting parameter.  In Section~\ref{sec:results}, we
compare these four models with the tortuosity of eroding porous
geometries.

@@

We showed in Sections~\ref{sec:DLPcomplex} and~\ref{sec:gradDLPcomplex}
that the velocity, shear stress, and vorticity of the double-layer
representation can all be written as the sum of Cauchy integrals and its
first two derivatives.  Therefore, we can accurately compute the
velocity, shear stress, and vorticity at any target location by using a
quadrature rule with a uniform error bound for Cauchy integrals and its
derivatives~\eqref{eqn:cauchy}. \citet{ioa-pap-per1991} developed
quadrature rules, that we call {\em Barycentric quadrature rules}, to
compute Cauchy integrals and their derivatives with an error bound that
is independent of $x \in \CC$.  Then,~\citet{bar-wu-vee2015} used these
quadrature rules to compute the Stokes double-layer potential
representation of the velocity~\eqref{eqn:velocityDLP}, and we briefly
summarize their method.  Then, we extend their work to compute the
second derivative so that the shear stress and vorticity can be computed
with a uniform error bound.

@@

To compute the shear stress and vorticity, we require a Barycentric
quadrature rule for $v''(x)$.  The derivation is largely based on the
work of~\citet{ioa-pap-per1991} (see equation (2.12)).  We start with
the second derivative of the Cauchy integral theorem
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y)}{(y-x)^3}\,dy - v''(x).
\end{align}
For the interior case, $x \in \Omega$, we use the identity
\begin{align}
  \frac{1}{2\upi i}\int_{\bd\Omega} \frac{1}{(y-x)^n}\, dy = 
  \left\{
    \begin{array}{ll}
      1, & n = 1, \\
      0, & n = 2,3,\ldots. \\
    \end{array}
  \right.
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y) - v''(x)(y-x)^2 - 2v(x) - 2(y-x)v'(x)}
      {(y-x)^3}\, dy.
\end{align}
This integrand is constructed so that it is bounded for all $x \in
\Omega$, and applying the trapezoid rule, we have
\begin{align}
  0 \approx  \sum_{j=1}^{N} 
      \frac{2v^{-}(y_j) - v''(x)(y_j-x)^2 - 2v(x) - 2(y_j-x)v'(x)}
      {(y_j-x)^3} w_j,
\end{align}
where the accuracy is independent of $x$.  Solving for $v''(x)$, the
Barycentric quadrature rule for the interior second derivative is
\begin{align}
  v''(x) &\approx \frac{2\sum\limits_{j=1}^N 
    \frac{v^{-}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{1}{y_j-x}w_j}, \quad x \in \Omega.
\end{align}
For the exterior case, $x \in \Omega^c$, we start with the identity
\begin{align}
\frac{1}{x-a} &= -\frac{1}{2\upi i}\int_{\bd\Omega} 
    \frac{(y-a)^{-1}}{y-x}\, dy. 
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
    \frac{2v^+(y) - 2v(x) - 2(y-x)v'(x) + (y-a)^{-1} (x-a) (y-x)^2 v''(x)}
    {(y-x)^3}.
\end{align}
As in the interior case, the integrand is chosen so that it is bounded
for all $x \in \Omega^c$.  Therefore, after applying the trapezoid rule
and solving for $v''(x)$, we have the Barycentric quadrature rule for
the exterior second derivative
\begin{align}
  v''(x) &\approx \frac{1}{x-a}\frac{2\sum\limits_{j=1}^N
    \frac{v^{+}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j}, 
    \quad x \in \Omega^c.
\end{align}
The quadrature rule for $v''(x)$ requires the values $v(x)$ and $v'(x)$,
and theses are computed using the quadrature rules in
equations~\eqref{eqn:BaryvInterior},~\eqref{eqn:BaryvExterior},~\eqref{eqn:BaryvprimeInterior},
and~\eqref{eqn:BaryvprimeExterior}. In Section~\ref{sec:results}, these
quadrature rules are used to form simulations of nearly-touching eroding
grains, and to study dynamics of the flow in regions arbitrarily close
to eroding grains.
